### 2024-12-03 other/datasheet/germany_datasheet.qmd

---
title: "Determinants of Soccer Players' Market Values - Germany"
author: John Zhang
thanks: "Code and data are available at: https://github.com/Clearsky21z/Player_Market_Value_Analysis"
date: today
date-format: long
format: pdf
number-sections: true
---

# **Motivation**

1. **For what purpose was the dataset created?**  
   To analyze determinants of soccer players’ market values, focusing on the Bundesliga. The goal is to enable robust insights for scouting, player valuation, and transfer market analysis.

2. **Who created the dataset and on behalf of which entity?**  
   The dataset was created by John Zhang as part of research into player market valuations.

3. **Who funded the creation of the dataset?**  
   The funding details are TBD.

4. **Any other comments?**  
   The dataset incorporates data from Transfermarkt, Stathead, FIFA, and Footballdatabase.

---

# **Composition**

1. **What do the instances that comprise the dataset represent?**  
   Each instance represents a Bundesliga player and their associated attributes, including market value, performance metrics, and contextual variables.

2. **How many instances are there in total?**  
   The dataset contains 385 player records.

3. **Is the dataset a sample or does it contain all possible instances?**  
   The dataset is a comprehensive sample of Bundesliga players for the 2023/24 season.

4. **What data does each instance consist of?**  
   Each instance includes variables such as player name, age, position, market value, club and national team rankings, goals, assists, and minutes played.

5. **Is there a label or target associated with each instance?**  
   The target variable is market value.

6. **Is any information missing from individual instances?**  
   No, the dataset is complete.

7. **Are relationships between individual instances made explicit?**  
   No explicit relationships are included.

8. **Are there recommended data splits?**  
   No, but splits for training and validation can be derived.

9. **Are there any errors, sources of noise, or redundancies?**  
   None identified.

10. **Is the dataset self-contained?**  
    Yes, the dataset is self-contained and integrates data from multiple reliable sources.

11. **Does the dataset contain confidential data?**  
    No confidential data is included.

12. **Does the dataset contain offensive or sensitive content?**  
    No offensive or sensitive content is included.

13. **Does the dataset identify any sub-populations?**  
    Sub-populations are identified by player positions.

14. **Is it possible to identify individuals directly or indirectly?**  
    Yes, players are identified by name.

15. **Does the dataset contain sensitive data?**  
    No sensitive data is included.

16. **Any other comments?**  
    None.

---

# **Collection Process**

1. **How was the data acquired?**  
   The data was scraped and collected from Transfermarkt, Stathead, FIFA, and Footballdatabase.

2. **What mechanisms or procedures were used to collect the data?**  
   A combination of web scraping and manual curation was used, scripts are available in the linked repository.

3. **What was the sampling strategy?**  
   Players were sampled based on their inclusion in the Bundesliga for the 2023/24 season.

4. **Who was involved in the data collection process?**  
   Employees for Transfermarkt, Stathead, FIFA, and Footballdatabase.

5. **Over what timeframe was the data collected?**  
   Data was collected during and after the 2023/24 Bundesliga season.

6. **Were ethical review processes conducted?**  
   TBD.

7. **Did you collect the data from the individuals directly or from third parties?**  
   Data was collected from third-party platforms.

8. **Were the individuals notified about the data collection?**  
   No notification was necessary as all data is publicly available.

9. **Did the individuals consent to data collection?**  
   Consent is not required for publicly available data.

10. **Was consent revocable?**  
    Not applicable.

11. **Was a data protection impact analysis conducted?**  
    TBD.

12. **Any other comments?**  
    None.

---

# **Preprocessing/Cleaning/Labeling**

1. **Was preprocessing done?**  
   Yes, the data underwent rigorous cleaning to ensure consistency and reliability.

2. **Was the raw data saved?**  
   Yes, raw data was retained for future reference.

3. **Is the preprocessing software available?**  
   Yes, preprocessing scripts are available in the linked repository.

4. **Any other comments?**  
   None.

---

# **Uses**

1. **Has the dataset been used for any tasks already?**  
   Yes, it was used in regression modeling to analyze market value determinants.

2. **Is there a repository linking papers or systems using this dataset?**  
   [GitHub Repository](https://github.com/Clearsky21z/Player_Market_Value_Analysis)

3. **What (other) tasks could the dataset be used for?**  
   Player scouting, valuation modeling, and market analysis.

4. **Does the dataset composition impact future uses?**  
   No significant limitations identified.

5. **Are there tasks for which the dataset should not be used?**  
   The dataset should not be used for non-sports-related analyses.

6. **Any other comments?**  
   None.

---

# **Distribution**

1. **Will the dataset be distributed to third parties?**  
   Yes, under an open-access license.

2. **How will the dataset be distributed?**  
   Via GitHub and linked repositories.

3. **When will the dataset be distributed?**  
   It is already available.

4. **Will the dataset be distributed under a license?**  
   Yes, an open-access license.

5. **Have third parties imposed restrictions?**  
   No.

6. **Do export controls or other regulations apply?**  
   No.

7. **Any other comments?**  
   None.

---

# **Maintenance**

1. **Who will support/host/maintain the dataset?**  
   John Zhang and his potential research team.

2. **How can the owner/curator/manager of the dataset be contacted?**  
   Contact details available in the linked GitHub repository.

3. **Is there an erratum?**  
   Not at this time.

4. **Will the dataset be updated?**  
   Future updates will include subsequent seasons.

5. **Are there retention limits?**  
   None.

6. **Will older versions of the dataset be supported?**  
   Yes, older versions will remain accessible.

7. **Is there a mechanism for extensions or contributions?**  
   Contributions are welcome via GitHub.

8. **Any other comments?**  
   None.

---

### **References**
Please refer to the paper and repository linked above for further information.